Tulai

### 2023-10-03 paper/paper.qmd
add pairwise tests for PCA results

>>>>

```{r}
#| label: tbl-permanova

# PerMANOVA - partitioning the euclidean distance matrix by species
# from https://stackoverflow.com/a/20267537/1036500
library(vegan)

pca_data_for_post_hoc_testing <-
pca_data %>%
  add_column(excavation_area)

# Is there are significant difference in the groups that
# we see in the PCA?
PerMANOVA_results <-
adonis2(pca_data[, 1:5] ~ excavation_area,
        data = pca_data_for_post_hoc_testing,
        method='eu')

PerMANOVA_results_inline <- 
  paste0("_F_(",
         PerMANOVA_results$Df[1],
         ", ",
         PerMANOVA_results$Df[2],
         ") = ",
         round(PerMANOVA_results$F[1], 3),
         " _p_ = ",
         PerMANOVA_results$`Pr(>F)`[1]
         )

# Which specific pairs of groups differ from each other?
# https://github.com/pmartinezarbizu/pairwiseAdonis
library(pairwiseAdonis)
pairwise.adonis2_results <-
pairwise.adonis2(pca_data[, 1:5] ~ excavation_area,
                data = pca_data_for_post_hoc_testing)

# turn this output into a nice table
pairwise.adonis2_results_tbl <- 
tibble(
  # names
  Pair =   c(names(pairwise.adonis2_results[2]),
             names(pairwise.adonis2_results[3]),
             names(pairwise.adonis2_results[4])),
  # F stat
       F = c(pairwise.adonis2_results[2][[1]]$F[1],
             pairwise.adonis2_results[3][[1]]$F[1],
             pairwise.adonis2_results[4][[1]]$F[1]), 
  # p-values
       p = c(pairwise.adonis2_results[2][[1]]$`Pr(>F)`[1],
             pairwise.adonis2_results[3][[1]]$`Pr(>F)`[1],
             pairwise.adonis2_results[4][[1]]$`Pr(>F)`[1])
) %>% 
  mutate(Pair = str_replace_all(Pair, "_", " "))

knitr::kable(pairwise.adonis2_results_tbl,
             caption = "Summary of pairwise post-hoc tests on the excavation areas")

A permutation MANOVA indicates that there is a significant difference in the lithic assemblage composition across the three excavation areas (`rPerMANOVA_results_inline`). We investigated this further using pairwise post-hoc tests and found that this significant difference is between TA and TP1, with the other two areas being statisically indistinguishable (@tbl-permanova )
